---
title: "A1-submission"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---






```{r}
library(tidyverse)
```

4a)
```{r}
passengers = read_csv("http://www.utsc.utoronto.ca/~butler/c32/airline_passengers.csv")

passengers
```


b) "the_date": date of the middle of the month which has "count""
```{r}
ggplot(passengers, aes(the_date,count))+geom_point()+geom_line()
```


c) There are 12 evenly (horizontally) spaced peaks approximately corresponding to each year. Overall, there is an increasing trend in terms of the number of passengers over the years.
There might be a pattern between number of passengers and different times of a year, by the ups and downs of the data. 


```{r}
ggplot(passengers, aes(mm,count))+geom_boxplot()
```


d) mm is a categorical variable. The graph of raw data is sorts mm by alphabetical order. Some work is needed to sort the months, if we want to see what happens as a year progresses. Nonetheless, by reading the median counts of each month, we can see the counts are higher during months mid-year of any year (Jun-Sep) and lower during end of year (Nov-Feb).


e) In part b), I inferred there might be a pattern associated between the number of passengers and different times of the year. This graph shows a similar trend, pointing out more people chose international air during certain seasons from the years 1949 to 1960.
