---
title: "A1-practice"
output: html_notebook
---


1 - JUICE
a)
```{r}
txt1="http://www.utsc.utoronto.ca/~butler/c32/ojuice.txt"
ojuice=read_delim(txt1, " ")
```
b)
```{r}
ojuice
```
c)
first line of txt file

d)
```{r}
ggplot(ojuice, aes(x=pectin,y=sweetness))+geom_point()+geom_smooth()
```

2 - SOAP
a) There are 27 rows:
```{r}
txt2 = "http://www.utsc.utoronto.ca/~butler/c32/soap.txt"
soap = read_delim(txt2, " ")

soap
```
b)
```{r}
ggplot(soap, aes(scrap)) + geom_histogram(bins=10)
```

c) By the graph of 10 bins, could describe it is approximately left-skewed (because of the large population falling on RHS), with exception of statistics of the group between approximately 240-270. Therefore in that sense, there is no significant pattern to be drawn other than the graph being right-heavy. Notable though would be the two relative peaks around 240-270 and 350-380. If the graph is looked at in other number of bins (e.g.7,9 or 14), it would appear bimodal.

d)
```{r}
ggplot(soap, aes(line, scrap)) + geom_boxplot()
```
e)
Yes, there is a difference as the median indicates so. Line b has a lower median, and most data fall into a lower overall range, meaning it produces less scrap than does line a. Most scrap production lies on the 270-370 range for both lines, which may indicate an overall pattern if given further data.

f)
```{r}
ggplot(soap, aes(speed, scrap))+geom_point()
```
g) There seems to be a positive correlation between the speed and amount of scrap produced.

.
.
.


