---
title: "A2-submission"
output:
  word_document: default
  html_notebook: default
---

```{r}
#8a)
library(tidyverse)
link="http://www.utsc.utoronto.ca/~butler/c32/car-speeds.csv"
speeds = read_csv(link)
speeds

max_speed = max(speeds$speed)
min_speed = min(speeds$speed)
count = max(speeds$car_no)
```

As we see, we have 1000 observed cars, having speeds within the range of [25,38]. Without knowing information about the stretch of road, this seems like appropriate information.


```{r}
#b)
nclass.Sturges(speeds$speed)
ggplot(speeds,aes(speed))+geom_histogram(bins=11)
```

c)
Most cars are close to normally distributed across various speeds between 25-38, with exceptions that i) there are over 300 cars (close to 1/3 of the total observed count) at 28 miles/hr, & ii) there are very few (total of less than 100) cars which go over 32miles/hr. 

d)
Based on the data from c, an appropriate speed limit would be of 32 miles per hour. Seeing how most of the cars go lower or at 32 miles per hour, it could imply that given the road condition,most drivers feel comfortable to pass safely at 32 miles/hr at the particular stretch of road.


```{r}
#9a) 
link2 = "http://www.utsc.utoronto.ca/~butler/c32/Backpack.csv"
backpacks=read_csv(link2)
backpacks
```
The name of the column containing "backpack weight (divided by) body weight" is "Ratio".


```{r}
#b)
nclass.Sturges(backpacks$Ratio)
ggplot(backpacks,aes(Ratio))+geom_histogram(bins=8)
```

```{r}
t.test(backpacks$Ratio, mu=0.10,alternative="less")
```
c) There is evidence that the mean  backpack-to-body-weight ratio is less than 0.10 by t-test. Since the p-value is significantly less than 0.05, we reject the null hypothesis that the mean is 0.10.



```{r}
#d)
t.test(backpacks$Ratio, conf.level = 0.9)
```

The 90% C.I. is [0.07104224,0.08320869] Meaning we are 90% confident that the ratio falls within [0.07104224,0.08320869].

e) Since the assumption for using a t-test is that the distribution is normal, t-test might not be appropriate. The graph for the t-test is right-skewed, therefore we could say it is non-normal. We can further check  with the Q-Q plot to determine if the distribution is normal.