---
title: "STAC32 A4Submission"
output:
  word_document: default
  html_notebook: default
---

```{r}
#9a
library(tidyverse)
file = "http://www.utsc.utoronto.ca/~butler/c32/nbapoints.csv"
score = read_csv(file)
score
```
We have date of the game and number of points for each game.

```{r}
#9b
ggplot(score, aes(sample=score$Points)) + stat_qq() + stat_qq_line()
```

c) There are a small number of outliers on either end (for they are away from the line), but most of the points scored generally follow a normal distribution. Since the data is generally normal, we infer that t-test is can be an appropriate choice of test. However with the highly deviated outliers, it can be said to be 'safer' to use a sign test.

```{r}
#9d
library(smmr)

#test against 183.2
sign_test(score,Points,183.2)
```

The sign test is already two-sided, which makes sense by the evenly distributed data points. Since the p-value is very low (<0.05), we can conclude to have strong evidence against the null hypothesis, so the average point is not 183.2.

```{r}
#10a
file2 <- "https://www.utsc.utoronto.ca/~butler/c32/salaryinc.txt"
salary = read_delim(file2," ")
salary
```
There are two columns for salaries in January 2016 and 2017 for each employee.

b)A match-pairs approach is better because we are comparing salaries of the same 20 employees with a before-and-after relationship. In order for a two-sample test to work, we would need to compare salaries of different employees at random. There would be more difficulties and variability implementing an alternative with two-sample.

```{r}
#10c
#want: Ha = diff > 0 only => want to reject H0 (diff = 0)
salary %>% mutate(diff=jan2017-jan2016) %>% ggplot(aes(sample=diff))+stat_qq()+stat_qq_line()
```

With a small sample size (n=20), there are a considerable number of outliers on the right. There are signs of the distribution being slightly right-skewed. With above said, the assumption for a matched-pair test fails.

```{r}
#10d
with(salary,t.test(jan2017,jan2016,paired=T))
```
With the small p-value, we can be confident to reject the null hypothesis and say the salary different is not 0.

e) From the output in d), we can see at "mean of the differences" that there is an average of 6.185 (dollar unit) increase for the two years. 

