---
title: "STAC51 A1Q5"
output:
  word_document: default
  html_notebook: default
---

5)
b)

```{r}
pi  <- 0.55 
pnaught <- 0.5
alpha <- 0.05
n <- 100

power_b <- 1-pnorm(((pnaught-pi)/(pi*(1-pi)/n)^0.5) + 2.33*(pnaught*pnaught/(pi*(1-pi))^0.5),0,1)
power_b

1-power_b

```
Comment: The power indicates there is 0.5658551 chance for a type II error. The score test will reject a false null hypothesis 0.4341449 chance at any given time.

c)

```{r}
x <- seq(0.4,0.7,0.005)

power_c <- 1-pnorm(((pnaught-x)/(x*(1-x)/n)^0.5) + 2.33*(pnaught*pnaught/(x*(1-x))^0.5),0,1)
plot(x,power_c, main="Power Function", xlab="pi", ylab="Power of x", type="l")
```

Comment: The power function seems to be monotonic, with both of its ends approaching extreme powers of 0 and 1. With a true value of > 0.65 (approximately), we are almost certain we can reject a false null hypothesis correctly.

d)
Find pi for power(π) = 0.5.

```{r}
x_d <- seq(0.5,0.6,0.0005)
power_d <- 1-pnorm(((pnaught-x_d)/(x_d*(1-x_d)/n)^0.5) + 2.33*(pnaught*pnaught/(x_d*(1-x_d))^0.5),0,1)

table <- cbind(x_d, power_d)
#show the subset that we want
subset(table, power_d > 0.495 & power_d < 0.51)

```

Answer: When power = 0.5, π is approximately 0.558.

e)

```{r}
n2 <- 200
n3 <- 300

plot(x,power_c, main="Power Function", xlab="pi", ylab="Power of x", type="l")
par(new=T)
power_e200 <- 1-pnorm(((pnaught-x)/(x*(1-x)/n2)^0.5) + 2.33*(pnaught*pnaught/(x*(1-x))^0.5),0,1)
plot(x,power_e200, main="Power Function", xlab="pi", ylab="Power of x", type="l", col="red")
par(new=T)
power_e300 <- 1-pnorm(((pnaught-x)/(x*(1-x)/n3)^0.5) + 2.33*(pnaught*pnaught/(x*(1-x))^0.5),0,1)
plot(x,power_e300, main="Power Function", xlab="pi", ylab="Power of x", type="l", col="green")
par(new=T)
```

Comment: The three curves show the same trends of starting at low power until approx. 0.5 and converge at around 1 as the curve grows. As n gets larger, the curve is steeper, meaning we gain power quickly (spend time at less number of values for pi).