---
title: "STAC51A2Q3"
output:
  word_document: default
  html_notebook: default
---
```{r}
#3a)
set.seed(101)
rn <- rbinom(2, 10, 0.4)
rn

```

```{r}
#3b)
n11 <- rn[1]
n21 <- rn[2]
n1plus <- 10
n2plus <- 10
ntable <- array(data = c(n11, n21, n1plus-n11, n2plus-n21), dim = c(2,2))
ntable
```



```{r}
#3c
ft <- fisher.test(ntable)
ft
#get p-value
ft$p.value
```
P-value large (>0.05) at 5% level significance. Cannot reject null hypothesis of independence between row and column vars.

```{r}
#3d)
#setup
set.seed(101)
pvals <- vector()
n1plus <- 10
n2plus <- 10

for (i in 1:10000) {
  #generate random numbers
  rn <- rbinom(2, 10, 0.4)
  #generate tables
  n11 <- rn[1]
  n21 <- rn[2]
  
  #fisher test on respective table
  pvals[i] <- fisher.test(array(data = c(n11, n21, n1plus-n11, n2plus-n21), dim = c(2,2)))$p.value
}
#the first 5 p-values
pvals[1:5]
#count the number of successful rejects for null hypothesis
rejects <- sum(pvals<0.05)
rejects
```
Null hypothesis is rejected 116 times out of 10000. So we're more confident to not reject the H0 of independence of the row and column variables.

```{r}
#setup
set.seed(101)
pvals <- vector()
n1plus <- 100
n2plus <- 100

for (i in 1:10000) {
  #generate random numbers
  rn <- rbinom(2, 100, 0.4)
  #generate tables
  n11 <- rn[1]
  n21 <- rn[2]
  
  #fisher test on respective table
  pvals[i] <- fisher.test(array(data = c(n11, n21, n1plus-n11, n2plus-n21), dim = c(2,2)))$p.value
}
#the first 5 p-values
pvals[1:5]
#count the number of successful rejects for null hypothesis
rejects <- sum(pvals<0.05)
rejects
```
Null hypothesis is rejected 354 times out of 10000. Again, we're more confident to not reject the H0 of independence of the row and column variables. This time a larger n, we got more successful rejects of H0.


